---
title: "iModEst_Figure4abc"
author: "Lauren Erdman"
date: "January 16, 2023"
output: html_document
---

```{r, include=FALSE}
### Required packages

library("ggplot2")
library("reshape")
library("RColorBrewer")
library("httr")

```


```{r, include=FALSE}

### Functions used below
read_remote_rds <- function(url) {
  # Make a GET request and save the response
  response <- GET(url)
 
  # Check if the request was successful
  stop_for_status(response)
 
  # Write the raw content to a temporary file and read it with readRDS
  temp_file <- tempfile(fileext = ".rds")
  writeBin(content(response), temp_file)
  data <- readRDS(temp_file)
 
  # Return the data
  return(data)
}


add.alpha <- function(col, alpha=1){
  if(missing(col))
    stop("Please provide a vector of colours.")
  apply(sapply(col, col2rgb)/255, 2, 
        function(x) 
          rgb(x[1], x[2], x[3], alpha=alpha))  
}

plot.r2.addition <- function(df,row.num,ylim=c(-1,1)){
  vars.removed <- matrix(nrow=0,ncol=3)
  colnames(vars.removed) <- c("var.added","base.model","r2.growth")
  
  for(k in rev(unique(nchar(colnames(df))))[1:(length(rev(unique(nchar(colnames(df)))))-1)]){
    mods <- colnames(df)[which(nchar(colnames(df)) == k)]
    mods.min.1 <- colnames(df)[which(nchar(colnames(df)) == (k-1))]
    
    for(i in 1:6){
      vars.removed <- rbind(vars.removed,cbind(rep(i,length(mods.min.1[grep(pattern = i,x = mods.min.1,invert = TRUE)])),mods.min.1[grep(pattern = i,x = mods.min.1,invert = TRUE)],unlist(df[row.num,mods[grep(pattern = i,x = mods)]] - df[row.num,mods.min.1[grep(pattern = i,x = mods.min.1,invert = TRUE)]])))
    }
  }
  
  # dim(vars.removed)
  # head(vars.removed)
  
  vars.removed <- as.data.frame(vars.removed)
  vars.removed$r2.growth <- as.numeric(as.character(vars.removed$r2.growth))
  vars.removed$var.added.f <- factor(vars.removed$var.added,levels=1:6,labels=c("miRNA","TF","lncRNA","methyl","CNV","SNP"))
  vars.removed$base.model <- as.character(vars.removed$base.model)
  # str(vars.removed)
  # head(vars.removed)
  
  ## DARKER = VARIABLE ADDED TO MODEL WITH FEWER PREDICTORS
  graph.list <- split(x = vars.removed$r2.growth,f = vars.removed$var.added.f)
  # sapply(1:)
  alpha.list <- split(x = 1-nchar(x = vars.removed$base.model)/10,f = vars.removed$var.added.f)
  
  # par(mfrow=c(1,1))
  # stripchart(split(x = vars.removed$r2.growth,f = vars.removed$var.added.f),method = "jitter",vertical = T,pch=19,col = add.alpha(col = "darkorchid",alpha = 0.7))
  stripchart(split(x = vars.removed$r2.growth,f = vars.removed$var.added.f),method = "jitter",vertical = T,pch=19,col = 'white',ylim=ylim,
             xlab = "",ylab = "",cex=2)
  for(k in 1:6){
    points(x = rep(k,length(graph.list[[k]])) + rnorm(n = length(graph.list[[k]]),0,0.1),y = graph.list[[k]],col = add.alpha(col = "blue",alpha.list[[k]]*2),pch=19,cex=2)
    points(x = k, y = df[row.num,paste0("glm_",k)],pch = 18,col = 'red',cex=2)
  }
  
  
}

plot.r2.addition.ZERO <- function(df,row.num,ylim=c(-1,1)){
  vars.removed <- matrix(nrow=0,ncol=3)
  colnames(vars.removed) <- c("var.added","base.model","r2.growth")
  
  df[df < 0] <- 0
  
  for(k in rev(unique(nchar(colnames(df))))[1:(length(rev(unique(nchar(colnames(df)))))-1)]){
    mods <- colnames(df)[which(nchar(colnames(df)) == k)]
    mods.min.1 <- colnames(df)[which(nchar(colnames(df)) == (k-1))]
    
    for(i in 1:6){
      vars.removed <- rbind(vars.removed,cbind(rep(i,length(mods.min.1[grep(pattern = i,x = mods.min.1,invert = TRUE)])),mods.min.1[grep(pattern = i,x = mods.min.1,invert = TRUE)],unlist(df[row.num,mods[grep(pattern = i,x = mods)]] - df[row.num,mods.min.1[grep(pattern = i,x = mods.min.1,invert = TRUE)]])))
    }
  }
  
  # dim(vars.removed)
  # head(vars.removed)
  
  vars.removed <- as.data.frame(vars.removed)
  vars.removed$r2.growth <- as.numeric(as.character(vars.removed$r2.growth))
  vars.removed$var.added.f <- factor(vars.removed$var.added,levels=1:6,labels=c("miRNA","TF","lncRNA","methyl","CNV","SNP"))
  vars.removed$base.model <- as.character(vars.removed$base.model)
  # str(vars.removed)
  # head(vars.removed)
  
  ## DARKER = VARIABLE ADDED TO MODEL WITH FEWER PREDICTORS
  graph.list <- split(x = vars.removed$r2.growth,f = vars.removed$var.added.f)
  # sapply(1:)
  alpha.list <- split(x = 1-nchar(x = vars.removed$base.model)/10,f = vars.removed$var.added.f)
  
  # par(mfrow=c(1,1))
  # stripchart(split(x = vars.removed$r2.growth,f = vars.removed$var.added.f),method = "jitter",vertical = T,pch=19,col = add.alpha(col = "darkorchid",alpha = 0.7))
  stripchart(split(x = vars.removed$r2.growth,f = vars.removed$var.added.f),method = "jitter",vertical = T,pch=19,col = 'white',ylim=ylim,
             xlab = "",ylab = "",cex=2)
  for(k in 1:6){
    points(x = rep(k,length(graph.list[[k]])) + rnorm(n = length(graph.list[[k]]),0,0.1),y = graph.list[[k]],col = add.alpha(col = "blue",alpha.list[[k]]*2),pch=19,cex=2)
    points(x = k, y = df[row.num,paste0("glm_",k)],pch = 18,col = 'red',cex=2)
  }
  
  
}

get.r2.addition.ZERO <- function(df,row.num,ylim=c(-1,1)){
  vars.removed <- matrix(nrow=0,ncol=4)
  colnames(vars.removed) <- c("var.added","base.model","r2.growth","mod.r2")
  
  df[df < 0] <- 0
  
  for(k in rev(unique(nchar(colnames(df))))[1:(length(rev(unique(nchar(colnames(df)))))-1)]){
    mods <- colnames(df)[which(nchar(colnames(df)) == k)]
    mods.min.1 <- colnames(df)[which(nchar(colnames(df)) == (k-1))]
    
    for(i in 1:6){
      vars.removed <- rbind(vars.removed,
                            cbind(rep(i,length(mods.min.1[grep(pattern = i,x = mods.min.1,invert = TRUE)])),
                                  mods.min.1[grep(pattern = i,x = mods.min.1,invert = TRUE)],
                                  unlist(df[row.num,mods[grep(pattern = i,x = mods)]] - 
                                           df[row.num,mods.min.1[grep(pattern = i,x = mods.min.1, invert = TRUE)]]),
                                  unlist(df[row.num,mods[grep(pattern = i,x = mods)]] - 
                                           rep(0,length(df[row.num,mods.min.1[grep(pattern = i,x = mods.min.1, invert = TRUE)]])))))
    }
  }
  
  # dim(vars.removed)
  # head(vars.removed)
  
  vars.removed <- as.data.frame(vars.removed)
  vars.removed$r2.growth <- as.numeric(as.character(vars.removed$r2.growth))
  vars.removed$mod.r2 <- as.numeric(as.character(vars.removed$mod.r2))
  vars.removed$var.added.f <- factor(vars.removed$var.added,levels=1:6,labels=c("miRNA","TF","lncRNA","methyl","CNV","SNP"))
  vars.removed$base.model <- as.character(vars.removed$base.model)
  vars.removed$gene <- rownames(df)[row.num]
  # str(vars.removed)
  # head(vars.removed)
  
  return(vars.removed) 
  
  ## DARKER = VARIABLE ADDED TO MODEL WITH FEWER PREDICTORS
  # graph.list <- split(x = vars.removed$r2.growth,f = vars.removed$var.added.f)
  # alpha.list <- split(x = 1-nchar(x = vars.removed$base.model)/10,f = vars.removed$var.added.f)
  
  # par(mfrow=c(1,1))
  # stripchart(split(x = vars.removed$r2.growth,f = vars.removed$var.added.f),method = "jitter",vertical = T,pch=19,col = add.alpha(col = "darkorchid",alpha = 0.7))
  
  ## STRIPCHART
  # stripchart(split(x = vars.removed$r2.growth,f = vars.removed$var.added.f),method = "jitter",vertical = T,pch=19,col = 'white',ylim=ylim,
  #            xlab = "",ylab = "",cex=2)
  # for(k in 1:6){
  #   points(x = rep(k,length(graph.list[[k]])) + rnorm(n = length(graph.list[[k]]),0,0.1),y = graph.list[[k]],col = add.alpha(col = "blue",alpha.list[[k]]*2),pch=19,cex=2)
  #   points(x = k, y = df[row.num,paste0("glm_",k)],pch = 18,col = 'red',cex=2)
  # }
  
  
}

get.mean.add.r2.by.reg <- function(r2.added.can.list){
  means.by.reg.list <- lapply(r2.added.can.list,function(x){
    sapply(1:6,function(i){
      mean(x$r2.growth[x$var.added == as.character(i)])
    })
  })
  return(means.by.reg.list)
}


# qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
# col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
# col_vector = col_vector[seq(1,length(col_vector),round(length(col_vector)/22))]

gg_color_hue <- function(n=21) {
  ## ggplot default rainbow colors from: https://stackoverflow.com/questions/8197559/emulate-ggplot2-default-color-palette 
  # hues = seq(15, 375, length = n + 1)
  
  hues = seq(25, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

```

##Set overall values
```{r}

tum.cols <- gg_color_hue(21)

reg.cols <- brewer.pal(n = 6,name = "Dark2")

cans <- "BLCA, BRCA, CESC, ESCA, HNSC, KIRC, KIRP, LGG, LIHC, LUAD, LUSC, PAAD, PCPG, PRAD, READ-COAD, SARC, SKCM, STAD, TGCT, THCA, UCEC"
tums <- unlist(strsplit(x = cans,split = ",\ "))
names(tum.cols) <- tums
tum.vec <- tums
cancers.of.int <- tums

r2_data_path = 'https://imodesttool.com/api/data/download/'


```

##Read in tumour data
### (Run initially to create R2 added values, commented out here)
```{r}

# tumour.r2.dfs <- list()
# gene.tum.details.df <- list()
# 
# k=1
# for(cancer in cancers.of.int){
#   
#   ## Reading in PRESS R2 results
#   tumour.r2.dfs[[k]] <- read_remote_rds(paste0(r2_data_path,'/full/performance_files/tumour_PRESS_R2/',cancer,"_0_tumor_r_square_1.rds")) 
#   # cancer <- gsub(pattern = "_.*$",replacement = "",x = tumour.file.in)
#   names(tumour.r2.dfs)[k] <- cancer
#   
#   ## Reading in cancer model details
#   ## Reading in cancer model details
#   summary_file = paste0(cancer,"_0_data_summary.rds")
#   
#   
#   ## Reading in cancer model details
#   gene.tum.details.df[[k]] <- read_remote_rds(paste0(r2_data_path,'/full/performance_files/summary_files/',summary_file))
#   names(gene.tum.details.df)[k] <- cancer
#   
#   k=k+1
# }
# 
# 
# tum.r2.dfs  <- tumour.r2.dfs
# tum.r2.dfs.sub  <- tumour.r2.dfs
# 
# ### MAKING OVERALL GENE SUMMARY
# num.reg.sum <- lapply(gene.tum.details.df,function(x){x$NUM_REGULATORS})
# overall.summary <- Reduce(rbind,num.reg.sum)
# overall.summary <- overall.summary[!duplicated(overall.summary$target_id),]


```

##Create lists of added variation explained and clusters of genes
### (This takes a long time to run so the files have been saved and this has been commented out uncomment to run and save new files)
```{r}

## 
##    CLUSTERING GENES -- only needs to be run once, output written to files
##

# library(cluster)

# cancers.of.int = c("BLCA","BRCA","CESC","ESCA","HNSC","KIRC","KIRP", "LIHC", "LGG","LUAD","LUSC","PAAD","PCPG","PRAD","READ-COAD","SARC","SKCM","STAD","TGCT","THCA","UCEC")

# gene.info.list = as.data.frame(gene.tum.details.df[[cancer]][[1]])

# gene.names.list <- lapply(t(tum.r2.dfs),colnames)
# gene_counts = unlist(lapply(gene.names.list,length))
# names(gene_counts) = names(tum.r2.dfs)
# write.csv(melt(gene_counts), file = "C:/Users/larun/Desktop/Modes/Data/gene_counts_20210506.csv")

# all_genes = Reduce(union, gene.names.list)

  ## only overlapping genes
# overlapping.genes <- na.omit(Reduce(f = function(x,y)intersect(x,y),gene.names.list))
# ol.genes.r2.dfs <- lapply(t(tum.r2.dfs),function(x){x[,overlapping.genes]})
  ## all genes
# ol.genes.r2.dfs <- lapply(t(tum.r2.dfs),function(x){x})
# names(ol.genes.r2.dfs) <- names(tum.r2.dfs)


# for(cancer in cancers.of.int[1:length(cancers.of.int)]){
#   if(length(grep(pattern = cancer,tumour.file.list)) == 1){
#     # i = 1
#     var.added.list <- list()
#     var.added.df <- data.frame(matrix(nrow=nrow(t(ol.genes.r2.dfs[[cancer]])),ncol=nrow(get.r2.addition.ZERO(df = t(ol.genes.r2.dfs[[cancer]]),row.num = 1))))
#     for(i in 1:nrow(t(ol.genes.r2.dfs[[cancer]]))){
#       var.added.df[i,] <- get.r2.addition.ZERO(df = t(ol.genes.r2.dfs[[cancer]]),row.num = i)$r2.growth
#       var.added.list[[i]] <- get.r2.addition.ZERO(df = t(ol.genes.r2.dfs[[cancer]]),row.num = i)
#     }
# 
#     # if(cancer == "THCA"){
#     rownames(var.added.df) <- row.names(t(ol.genes.r2.dfs[[cancer]]))
#     names(var.added.list)  <- row.names(t(ol.genes.r2.dfs[[cancer]]))
#       
#     # }
#     #     
#     # ## SHOULD BE THE SAME ACROSS CANCERS, HAD ISSUE WITH THCA SO JUST LEAVING IT WITH HNSC
#     # rownames(var.added.df) <- gene.tum.details.df[[cancer]][[1]]$gene_name[match(row.names(t(ol.genes.r2.dfs[[cancer]])),gene.tum.details.df[[cancer]][[1]]$target_id)]
#     # names(var.added.list)  <- gene.tum.details.df[[cancer]][[1]]$gene_name[match(row.names(t(ol.genes.r2.dfs[[cancer]])),gene.tum.details.df[[cancer]][[1]]$target_id)]
# #
# #     # str(var.added.list)
#     # dist.out <- dist(var.added.df)
# #     # str(dist.out)
#     # hclust.out <- hclust(dist.out)
# #
#     # clusts <- sapply(2:100,function(i){cutree(hclust.out,i)})
#     # sil.means <- sapply(2:100,function(i){test <- cutree(tree = hclust.out,k = i) ; sil.out <- silhouette(x = test,dist = dist.out) ; return(mean(sil.out[,3]))})
#     # sil.sds <- sapply(2:100,function(i){test <- cutree(tree = hclust.out,k = i) ; sil.out <- silhouette(x = test,dist = dist.out) ; return(sd(sil.out[,3]))})
# #
#     saveRDS(var.added.list,paste0("C:/Users/larun/Desktop/Modes/Data/Clustering-output/",cancer,"-var-added-list-allgenes.rds"))
#     saveRDS(var.added.df,paste0("C:/Users/larun/Desktop/Modes/Data/Clustering-output/",cancer,"-var-added-df-allgenes.rds"))
#     # saveRDS(dist.out,paste0("C:/Users/larun/Desktop/Modes/Data/Clustering-output/",cancer,"-dist-mat-20can.rds"))
#     # saveRDS(hclust.out,paste0("C:/Users/larun/Desktop/Modes/Data/Clustering-output/",cancer,"-hclust-20can.rds"))
#     # saveRDS(sil.means,paste0("C:/Users/larun/Desktop/Modes/Data/Clustering-output/",cancer,"-sil-means-20can.rds"))
#     # saveRDS(sil.sds,paste0("C:/Users/larun/Desktop/Modes/Data/Clustering-output/",cancer,"-sil-sds-20can.rds"))
#     # saveRDS(clusts,paste0("C:/Users/larun/Desktop/Modes/Data/Clustering-output/",cancer,"-clusts-20can.rds"))
# #
# #
#     cat(paste0("Cancer: ",cancer," complete. \n"))
# 
#   }
# }
# 

```

##Prep data for overall tumour graph
```{r}

r2.added.list = read_remote_rds(paste0(r2_data_path, "/full/singleton_files/r2addedlist_21can_allgenes_renamed_20220204.rds"))

ol.genes = Reduce(intersect,lapply(r2.added.list,names))
# length(ol.genes)

r2.added.list = lapply(r2.added.list,function(x)x[ol.genes])

## OVERLAPPING GENES ONLY
for(cancer in names(r2.added.list)){
  names(r2.added.list[[cancer]]) = ol.genes
}


r2.added.means.list <- lapply(r2.added.list,get.mean.add.r2.by.reg)

reg.means.df.by.can <- lapply(r2.added.means.list,function(x){
  bound.fields <- data.frame(Reduce(rbind,x))
  names(bound.fields) <- c("miRNA","TF","lncRNA","methyl","CNV","SNP")
  return(bound.fields)
  })

for(i in 1:length(reg.means.df.by.can)){
  rownames(reg.means.df.by.can[[i]]) <- names(r2.added.list[[1]])
}

# head(reg.means.df.by.can[[1]])
# lapply(reg.means.df.by.can,dim)

mydf <- reg.means.df.by.can[[1]]
mydf$cancer <- names(reg.means.df.by.can)[1]

for(i in 2:length(reg.means.df.by.can)){
  df.alone <- reg.means.df.by.can[[i]]
  df.alone$cancer <- names(reg.means.df.by.can)[i]
  
  mydf <- rbind(mydf,df.alone)  
}

# head(mydf)
# table(mydf$cancer)
# length(table(mydf$cancer))


```




##Prep data for gene graph
```{r}

mygene = "CRYBB2"

gene.list <- lapply(r2.added.list,function(x){
  if(mygene %in% names(x)){
    newdf <- x[[mygene]]
    rownames(newdf) <- 1:nrow(newdf)
    return(newdf)
  } else{
    return(NA)
  }
})

full.gene.df <- Reduce(rbind,gene.list[names(gene.list[is.na(gene.list) == FALSE])])
gene_spec_cancers = names(gene.list[is.na(gene.list) == FALSE])
full.gene.df$cancer <- rep(gene_spec_cancers,each = nrow(gene.list[is.na(gene.list) == FALSE][[1]]))


```

##CRYBB2 Fig 3c all miRNA models
```{r}

full.gene.df.sub = full.gene.df[full.gene.df$var.added.f == "SNP",]

ggplot(full.gene.df,aes(x = var.added.f, y = r2.growth,fill = cancer)) + 
  geom_boxplot() + xlab("Regulator") + ylab("R2 added") + ylim(c(-1,1)) + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold")) + 
  scale_fill_manual(values = tum.cols) + ggtitle(mygene)

# str(full.gene.df)


```

##CRYBB2 Fig 4b all SNP models with methyl
```{r}

cond.var = "4"
cond.var2 = ""
cond.var3 = ""

cond.gene.df.mlt <- full.gene.df[intersect(intersect(grep(cond.var,full.gene.df$base.model),grep(cond.var2,full.gene.df$base.model)),grep(cond.var3,full.gene.df$base.model)),]

cond.gene.df.mlt.sub = cond.gene.df.mlt[cond.gene.df.mlt$var.added.f == "SNP",]

ggplot(cond.gene.df.mlt.sub,aes(x = var.added.f, y = r2.growth,fill = cancer)) + 
  geom_boxplot() + xlab("Regulator") + ylab("R2 added") + ylim(c(-1,1)) + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold")) + 
  scale_fill_manual(values = tum.cols)

```

##CRYBB2 Fig 3c all SNP models no methyl
```{r}

cond.var = "4"
cond.var2 = ""
cond.var3 = ""

uncond.gene.df.mlt <- full.gene.df[grep(cond.var,full.gene.df$base.model,invert = TRUE),]

uncond.gene.df.mlt.sub = uncond.gene.df.mlt[uncond.gene.df.mlt$var.added.f == "SNP",]

ggplot(uncond.gene.df.mlt.sub,aes(x = var.added.f, y = r2.growth,fill = cancer)) + 
  geom_boxplot() + xlab("Regulator") + ylab("R2 added") + ylim(c(-1,1)) + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold")) + 
  scale_fill_manual(values = tum.cols)
                                
```

